---
title: "Behavioral analysis for the AX-CPT E-Prime log files"
author: "Qingkai Zhu"
date: "2023-06-18"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Behavioral analysis for the AX-CPT E-Prime log files

```{r pressure, echo=FALSE}
library(plyr)
library(dplyr)
library(readxl)
library(ggplot2)
library(tidyr)

rm(list = ls())

# Function to add all files in directory to one df
aggr_dat <- function (files, condition, aggreted_dat, all_dat){
  
  for (file in files){
    dat <- select(read_excel(paste0(data_dir, "/",condition, "/", file)), Subject, Name, Trial, cue, probe, probe.RT, probe.ACC)
    subj_name <- strsplit(file, "/")[[1]][2]
    # Create 'isOutlierTrial' column, initially setting all to 0
    dat$isOutlierTrial <- 0
    
    # Remove trials if RT is more than 3 SD from the mean
    correct_dat <- dat[dat$probe.ACC == 1,]
    lower_threshold <- mean(correct_dat$probe.RT) - 3*sd(correct_dat$probe.RT)
    upper_threshold <- mean(correct_dat$probe.RT) + 3*sd(correct_dat$probe.RT)

    dat$isOutlierTrial[dat$probe.ACC == 1 & (dat$probe.RT < lower_threshold | dat$probe.RT > upper_threshold)] <- 1

    dat$group <- condition
    
    # Average trials
    subj_aggr <- ddply(dat, c("cue", "probe", "Subject"), summarise,
                       conACC = mean(probe.ACC),
                       conRT = mean(probe.RT)
    )
    subj_aggr$cond <- condition
    # updated
    subj_aggr$meanRT = mean(dat$probe.RT)
    subj_aggr$meanACC = mean(dat$probe.ACC)
    subj_aggr$name = subj_name
    aggreted_dat <- rbind(aggreted_dat, subj_aggr)
    all_dat <- rbind(all_dat, dat)
  }
  return (list(aggreted_dat = aggreted_dat, all_dat = all_dat))
}

data_dir <- "C:/Users/chuch/Documents/GitHub/fyrp/data"
conditions <- c("LTA", "MTA", "HTA")

lta_files <- list.files(paste0(data_dir, "/LTA"), pattern="*/*.xlsx", recursive = T)
mta_files <- list.files(paste0(data_dir, "/MTA"), pattern="*/*.xlsx", recursive = T)
hta_files <- list.files(paste0(data_dir, "/HTA"), pattern="*/*.xlsx", recursive = T)

# Aggregate all anxiety groups into one df
cond_aggr <- data.frame()
all_dat <- data.frame()
results <- aggr_dat(lta_files, conditions[1], cond_aggr, all_dat)
cond_aggr <- results$aggreted_dat
all_dat <- results$all_dat

results <- aggr_dat(mta_files, conditions[2], cond_aggr, all_dat)
cond_aggr <- results$aggreted_dat
all_dat <- results$all_dat

results <- aggr_dat(hta_files, conditions[3], cond_aggr, all_dat)
cond_aggr <- results$aggreted_dat
all_dat <- results$all_dat

cond_aggr$cond <- factor(cond_aggr$cond, levels=c("LTA", "MTA", "HTA"))
cond_aggr$pair <- paste0(cond_aggr$cue, cond_aggr$probe)
all_dat$pair <- paste0(all_dat$cue, all_dat$probe)
```

```{r}
group_aggr <- ddply(all_dat, c("cue", "probe", "group"), summarise,
                   conACC = mean(probe.ACC), sderrACC = sd(probe.ACC) / sqrt(length(probe.ACC)),
                   conRT = mean(probe.RT), sderrRT = sd(probe.RT) / sqrt(length(probe.RT))
)

group_aggr

# Accuracy plot
ggplot(group_aggr, aes(x = interaction(cue, probe), y = conACC, fill = group)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = conACC - sderrACC, ymax = conACC + sderrACC), width = 0.2,
                position = position_dodge(.9)) +
  labs(x = "Cue-Probe Pair", y = "Mean Accuracy", fill = "Group") +
  theme_minimal() +
  theme(text = element_text(size = 12)) +
  scale_x_discrete(labels = c("A.X" = "A-X", "A.Y" = "A-Y", "B.X" = "B-X", "B.Y" = "B-Y"))  # Updated labels

# RT plot
ggplot(group_aggr, aes(x = interaction(cue, probe), y = conRT, fill = group)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = conRT - sderrRT, ymax = conRT + sderrRT), width = 0.2,
                position = position_dodge(.9)) +
  labs(x = "Cue-Probe Pair", y = "Mean Reaction Time", fill = "Group") +
  theme_minimal() +
  theme(text = element_text(size = 12)) +
  scale_x_discrete(labels = c("A.X" = "A-X", "A.Y" = "A-Y", "B.X" = "B-X", "B.Y" = "B-Y"))  # Updated labels

```

# Outlier Trials

```{r}
# Summarise data for group
outlier_summary_group <- all_dat %>% 
  filter(isOutlierTrial == 1) %>%  # Select only outlier trials
  group_by(group) %>%  # Group by 'group'
  summarise(count = n())  # Count number of outliers in each group
# Add proportion
outlier_summary_group <- outlier_summary_group %>% 
  mutate(prop = count / sum(count))  # Calculate proportion of total for each group
# Add labels to data for group
outlier_summary_group <- outlier_summary_group %>% 
  mutate(label = paste(count, "(", scales::percent(prop), ")"))
# Create pie chart with labels
ggplot(outlier_summary_group, aes(x = "", y = prop, fill = group)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar("y", start = 0) +
  theme_void() +
  theme(legend.position = "right") +
  guides(fill = guide_legend(title = "Group")) +
  labs(title = "Distribution of Outlier Trials by Group",
       subtitle = "Proportion of total outlier trials") +
  geom_text(aes(label = label), position = position_stack(vjust = 0.5))

# Summarise data for pair
outlier_summary_pair <- all_dat %>% 
  filter(isOutlierTrial == 1) %>%  # Select only outlier trials
  group_by(pair) %>%  # Group by 'pair'
  summarise(count = n())  # Count number of outliers in each pair
# Add proportion
outlier_summary_pair <- outlier_summary_pair %>% 
  mutate(prop = count / sum(count))  # Calculate proportion of total for each pair
# Add labels to data for pair
outlier_summary_pair <- outlier_summary_pair %>% 
  mutate(label = paste(count, "(", scales::percent(prop), ")"))
# Create pie chart with labels
ggplot(outlier_summary_pair, aes(x = "", y = prop, fill = pair)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar("y", start = 0) +
  theme_void() +
  theme(legend.position = "right") +
  guides(fill = guide_legend(title = "Pair")) +
  labs(title = "Distribution of Outlier Trials by Cue-Probe Pair",
       subtitle = "Proportion of total outlier trials") +
  geom_text(aes(label = label), position = position_stack(vjust = 0.5))

# Filter only the outlier trials
outlier_trials <- all_dat %>% filter(isOutlierTrial == 1)

# Get the range of RT for each pair
rt_ranges <- outlier_trials %>%
  group_by(pair) %>%
  summarise(min_rt = min(probe.RT, na.rm = TRUE), max_rt = max(probe.RT, na.rm = TRUE))

# Print the ranges
print(rt_ranges)

# Create a boxplot of RTs for each pair
ggplot(outlier_trials, aes(x = pair, y = probe.RT)) +
  geom_boxplot() +
  labs(x = "Cue-Probe Pair", y = "Response Time (RT)", 
       title = "Boxplot of Response Times for Outlier Trials by Cue-Probe Pair")

# Create a scatter plot of RTs for each pair
ggplot(outlier_trials, aes(x = pair, y = probe.RT)) +
  geom_point() +
  labs(x = "Cue-Probe Pair", y = "Response Time (RT)", 
       title = "Scatterplot of Response Times for Outlier Trials by Cue-Probe Pair")

```
# Outlier Subjects
There are five subjects who have less than 60% accuracy on BY trials, and correspondingly they also have low accuracy on BX trials.

- "Subject removed: Ganga BY_ACC: 0.364583333333333 BX_ACC: 0.427083333333333 AX_ACC: 0.988095238095238 AY_ACC: 0.927083333333333 group: LTA all acc: 0.863541666666667"
- "Subject removed: jayant BY_ACC: 0.510416666666667 BX_ACC: 0.5 AX_ACC: 0.979166666666667 AY_ACC: 0.729166666666667 group: MTA all acc: 0.859375"
- "Subject removed: kamal BY_ACC: 0.583333333333333 BX_ACC: 0.5625 AX_ACC: 0.96577380952381 AY_ACC: 0.65625 group: MTA all acc: 0.85625"
- "Subject removed: sonali BY_ACC: 0.59375 BX_ACC: 0.59375 AX_ACC: 0.793154761904762 AY_ACC: 0.666666666666667 group: MTA all acc: 0.740625"
- "Subject removed: aarav BY_ACC: 0.572916666666667 BX_ACC: 0.541666666666667 AX_ACC: 0.992559523809524 AY_ACC: 0.614583333333333 group: HTA all acc: 0.867708333333333"

```{r}
library(tidyr)
# Filter and summarise data
low_acc_summary <- cond_aggr %>%
  filter(conACC < 0.6) %>%
  group_by(pair) %>%
  summarise(count = n())

# Add proportion
low_acc_summary <- low_acc_summary %>%
  mutate(prop = count / sum(count),
         label = paste(count, "(", scales::percent(prop), ")"))

# Create pie chart
ggplot(low_acc_summary, aes(x = "", y = prop, fill = pair)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar("y", start = 0) +
  theme_void() +
  theme(legend.position = "right") +
  guides(fill = guide_legend(title = "Pair")) +
  labs(title = "Distribution of Low Accuracy Subject by Cue-Probe Pair",
       subtitle = "Proportion of total low accuracy trials") +
  geom_text(aes(label = label), position = position_stack(vjust = 0.5))

# Filter cond_aggr
low_acc_cond_aggr <- cond_aggr %>%
  filter(conACC < 0.6)

# Join with group_aggr
comparison_df <- left_join(low_acc_cond_aggr, group_aggr, by = c("cue" = "cue", "probe" = "probe", "cond" = "group"))

# View comparison_df
head(comparison_df)


ggplot(comparison_df, aes(x = conRT.x, y = conRT.y)) +
  geom_point() +
  geom_abline(intercept = 0, slope = 1, color = "red") +
  labs(x = "conRT (low_acc_cond_aggr)", y = "conRT (group_aggr)",
       title = "Comparison of conRT in low_acc_cond_aggr and group_aggr")

# Reformat data for easier plotting
comparison_df_melt <- comparison_df %>% 
  pivot_longer(c(conRT.x, conRT.y), names_to = "source", values_to = "conRT")

# Adjust source labels for clarity
comparison_df_melt$source <- ifelse(comparison_df_melt$source == "conRT.x", "low_acc_cond_aggr", "group_aggr")

# Plot
ggplot(comparison_df_melt, aes(x = pair, y = conRT, fill = source)) +
  geom_bar(stat = "identity", position = "dodge") +
  facet_wrap(~Subject) +
  labs(x = "Pair", y = "conRT", fill = "Source",
       title = "Comparison of conRT in low_acc_cond_aggr and group_aggr",
       subtitle = "Facetted by Subject") +
  theme_minimal()

# Summarise data
low_acc_summary_group <- low_acc_cond_aggr %>%
  group_by(cond) %>%  # Group by 'cond'
  summarise(count = n())  # Count number of rows in each condition

# Calculate proportion
low_acc_summary_group <- low_acc_summary_group %>% 
  mutate(prop = count / sum(count))  # Calculate proportion of total for each condition

# Create pie chart
ggplot(low_acc_summary_group, aes(x = "", y = prop, fill = cond)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar("y", start = 0) +
  theme_void() +
  theme(legend.position = "right") +
  guides(fill = guide_legend(title = "Condition")) +
  labs(title = "Distribution of Low Accuracy Trials by Condition",
       subtitle = "Proportion of total low accuracy trials")

```